---
title: "Empirical Research"
author: "Adham Rishmawi"
format: html
editor: visual
---

## Setting up packages and knitting options

```{r,setup,include=FALSE}
# load packages that are going to be used
require(tidyverse)   # this loads mosaic, ggformula, etc. too
require(ggformula)
require(mosaic)
library("glmmTMB")
library(lme4)
library(ggfortify)  # load ggfortify package for plotting functions
library(ggplot2)  # load ggplot2 package for theme customization
library(MuMIn)
library(mgcv)
library(DHARMa)
library(ggeffects)

# Some customization.  You can alter or delete as desired (if you know what you are doing).

theme_set(theme_bw(base_size=12))     # change theme for ggplot2/ggformula

knitr::opts_chunk$set(
  echo = TRUE,      # for Research, always show R code (this is the default)
  tidy = FALSE,     # display code as typed (rather than reformatted)
  size = "small",   # slightly smaller font for code
  message = FALSE, warning = FALSE) # don't print warnings or messages in compiled document. So you MUST check them in RStudio!
```

## Loading in data from source

```{r}
data1 <- read.csv("https://sldr.netlify.app/data/sustainable-livelihoods.csv")|>
  drop_na(Increased_Yield) |>
  drop_na(Increased_Knowledge)|>
  drop_na(Food_Ran_Out)|>
  arrange(Partner)

```

-   We dropped the NA instances from the main predictors and response variable we would like to use!

## Economic question

The economic question was derived out of the description of the data-set I found:

Q. **do the programs affect the individual's intellectuality and yield where it makes sure that there is less food deprivation occurring more often. (is there a association between food_ran_out and increased_knowledge and increased_yield?)**

**Data Description:**

World Renew implemented Sustainable Livelihoods, a five-year project in Bangladesh, Honduras, Mali, Mozambique, and Tanzania to enhance livelihood security for vulnerable households. The Sustainable Livelihoods program worked to build the adaptive capacity of individual households, as well as communities, to manage climate change risks. Poverty in the communities where the program operated is manifested by poor health, low incomes, food insecurity, landlessness, illiteracy and underemployment. Small-scale farmers are challenged by declining soil fertility, lack of secure land tenure, erratic weather and lack of access to credit and inputs. These conditions lead to food shortages and poor nutrition and health. In addition, poor urban households lack skills for employment and financing for small enterprises.

About 80% of the program participants were subsistence farmers who depend on rain-fed agriculture for their livelihoods. The other 20% of program participants belonged to poor households in urban areas, many of them living in slums where living conditions are crowded and unsanitary.

Participants had access to one or more training opportunities designed to promote or improve financial savings, leadership, literacy, and agricultural practices.

World Renew collected data to measure the impact of the Sustainable Livelihoods program including variables:

-   `Participant_ID`, unique ID for each participant

-   `Country`

-   `Partner` organization

-   `Gender` of participant

-   `Town` where the person lives

-   `District` where the person lives

-   `Age_Group` of the person completing the survey

-   `Savings_Program`, whether the person participated in the financial savings program

-   `Leadership_Program`, whether the person participated in the leadership training program

-   `Agriculture_Program`, whether the person participated in the sustainable agriculture program

-   `Literacy_Program`, whether the person participated in the literacy program

-   `n_Programs`, number of programs the person participated in

-   `FFS`, food frequency score (a measure of food security)

-   `FDS`, food diversity score (a measure of food security and diverse diet)

-   `Months_Enough_Food`, months (out of the 12 in the past year) where the person had enough food

-   `Months_Insufficient_Food`, months (out of the 12 in the past year) where the person *did not* have enough food

-   `Business_Plan`, whether the person had a business plan

-   `Management_Confidence`, whether the person had confidence in their management skills

-   `Sustainable_Ag_Practices`, whether the person practiced sustainable agriculture

-   `Minimum_Tillage`, whether the person used minimum tillage (a sustainable farming practice)

-   `Soil_Covered`, whether the person kept soil covered (a sustainable farming practice)

-   `Crop_Rotation`, whether the person rotated crops (a sustainable farming practice)

-   `All_Sustainable_Practices`, whether the person did all three sustainable farming practices: minimum tillage, soil covered, and crop rotation

-   `Increased_Yield`, whether the person had had increased farming yield after program participation

-   `Drought_Disease_Resistance`, whether the person was growing drought- or disease-resistant seeds

-   `Sustainable_Useful`, whether the person thought sustainable farming practices were useful

-   `Literacy_Helped`, whether the person thought the literacy program helped them

-   `Increased_Knowledge`, whether the person said their knowledge had increased

-   `Food_Ran_Out`, whether or not the person ran out of food during the study period

-   `Vegetables`, whether the person, others in the family, or no one ate vegetables

-   `Fruits`, whether the person, others in the family, or no one ate fruits

**Data reference/other conducted studies:**

-   Link to UN Sustainability development and Promoting Development Cooperation [here.](www.un.org/en/ecosoc/docs/pdfs/fina_08-45773.pdf)

-   link to a statement about the USA contribution to this type of initiative [here.](https://www.usaid.gov/news-information/press-releases/may-12-2022-usaid-launches-new-205-million-climate-change-and-environmental)

    ## 

    ## Is Economic question well motivated?

-   This question has not been proposed yet directly and is worth asking because the initiative will indicate to the provider of this program a factor to see if it is worthwhile.

-   The links provided above have done studies that are quite different but this question will focus more on the intellectuality increase and if that decreases the amount of food deprivation in these countries.

-   Whether this question indicates to us true or not it will give us a right step into indicating the more precise question we should be asking.

    ## Predictor choices for model

-   `Increased_Knowledge` and `Increased_Yield`

    -   Justification: This i thought would be a great predictor to pair up with increase_yield because they are both binary options which we can study instances where both are either 1 or zeros and see how they influence my response variable.

-   `Age_Group`

    -   Justification: this an additional predictor that i thought would influence my response variable because certain ages might be a contributor to the reason why they ran out of food or the reason their knowledge changed differently so I included it.

-   `Gender`

    -   Justification: this was a included because I believe cultural aspects and circumstances would alter the rates at which people ran out of food and learned.

-   `(1|Partner)`

    -   Justification: the Random effects I based it on The Partner because i believed it would be a bad traditional predictor because all the Organizations included have the same goal so theoretical it should be a bad predictor. It also fits the criteria we said in class that it must be a location, time, or identity which it is. I did not get any convergence problems so that's a good identification that it is a good RE effect.

-   All other columns and variables:

    -   Justification: I wanted to stay within the rule of thumb for this test and make sure that my predictors did not go beyond the threshold. Other predictors would have played into distorting the correlation between my response variable and predictors.

    ## Response variable Justification

-   the reason I choose `Food_Ran_Out` is because I want to do binary evaluation because this variable is either no or yes. I had other options such as `Months_insufficient_Food` but I thought this response variable would be more suitable because it is whether they completely ran out or not which is more simpler.

    ## Operationalization of each of the variables

-   the degree of measurement for our response variable is whether the people ran out of food more frequently less when we saw an increase in knowledge from these programs.

-   we will also tie an interaction with `Increased_Yield` and `Increased_Knowledge` to see if we can observe a corresponding reaction from both. (indicating if we see an increase in knowledge that in turn, results in a increased crop yield)

-   The quality of this data is derived from a established source so it has trust-able factors which means the extractions from it are tangible.

    ## Motivation to justify Research question

here are examples of researches that did similar studies under different constraints:

-   **The Impact of Livelihood Assets on the Food Security of Farmers in Southern Iran during the COVID-19 Pandemic.** The authors review a range of case studies and program evaluations from around the world to assess the impact of sustainable livelihoods programs on key indicators of food security, such as access to food, dietary diversity, and food utilization. [Link](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8156269/)

-   **Enhancing Agricultural Productivity and Sustainability through Sustainable Livelihoods Programs: Evidence from Ghana.** This study by Mensah and colleagues examines the impact of a sustainable livelihoods program on agricultural productivity and sustainability in Ghana. The authors use a randomized controlled trial design to assess the effectiveness of the program in improving farmers' access to inputs, promoting sustainable farming practices, and increasing yields. [Link](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6740034/)

-   **An Assessment of Poverty Alleviation Measures and Sustainable Livelihood Capability of Farm Households in Rural China: A Sustainable Livelihood Approach.** This study by Kumar and colleagues explores the relationship between sustainable livelihoods programs and knowledge acquisition in rural India. The authors use a mixed-methods approach to assess the impact of the program on farmers' knowledge of sustainable farming practices and their ability to adapt to climate change risks. [Link](https://www.mdpi.com/2077-0472/11/12/1230)

## Quantitative research

#### Initial demonstrating graphs

```{r}
int_graph1 <- gf_boxplot(Food_Ran_Out ~ Increased_Knowledge | Increased_Yield ,  data = data1) %>%
  gf_jitter(alpha = 0.3, width = 0.2) %>%
   gf_labs(subtitle = "Initial graph elaborating assiocation between knowledge and the decrease of no food",
          title = "Assiocation Graph for interaction",
          x = " Increase of Knowledge",
          y = "Did they have food?", 
           )

int_graph1
```

```{r}
int_graph2 <- gf_boxplot(Food_Ran_Out ~ Increased_Yield | Increased_Knowledge ,  data = data1) %>%
  gf_jitter(alpha = 0.3, width = 0.2) %>%
   gf_labs(subtitle = "Initial graph elaborating assiocation between Yield and the decrease of no food",
          title = "Assiocation Graph for interaction",
          x = " Increase of Food yield",
          y = "Did they have food?", 
           )

int_graph2
```

**Returning to the question we have in mind :**

do the programs affect the individual's intellectuality and yield where it makes sure that there is less food deprivation occurring more often.

ANSWER: Most definitely yes because we can observe an increase or a difference in the amount of people who had food and did/didnt take the program. I believe this makes my research question a topic worth going into and evaluating. We could conclude hesitantly so far that the programs have caused an increase in knowledge and indirectly influence our yield which has caused a effect in the amount of food deprived people.

#### Our Model

```{r}
model_fitRan <- glmmTMB(factor(Food_Ran_Out) ~ Increased_Knowledge*Increased_Yield + Age_Group + Gender +
                  (1|Partner), 
                 data = data1, 
                 family = binomial(link = 'logit'),
                 na.action = 'na.fail',
                 REML = FALSE)


msummary(model_fitRan)
```

**summary of model:**

-   **Model:** glmmTMB is form of regression which uses a mixed-effects linear model which is great for binomial regression. It is also used for random effects to which the parameter `Partner` would be our random effects.

-   **Interaction term:** we established an interaction predictor between `Increased_Yield` \* `Increased_Knowledge` so that is elaborated in this model

-   **na.action and na.fail:** will fail the model if there are any missing data

-   **Reml:** is set to **`FALSE`**, which means that the model is fitted using maximum likelihood estimation, rather than restricted maximum likelihood estimation.

    **Results of model:**

-   **Rationale =** The logic for the predictors still remains the same. However, the Random effects I based it on The Partner because i believed it would be a bad traditional predictor because all the Organizations included have the same goal so theoretical it should be a bad predictor. It also fits the criteria we said in class that it must be a location, time, or identity which it is. I did not get any convergence problems so that's a good identification that it is a good RE effect. However, we have a high AIC and BIC relative to our observation so we should continue testing our model to see if there is correlation.

-   **SUMMARY EXPLANATION**

    Random effect variance estimate was overall good showing that the inclusion of partner is correct based on the STd. dev.

    `Partner (intercept) 1.669 Variance 1.292 std. Dev`

    `number of obs: 188, groups: Partner, 10`

    ## Assessments

    **Dredge**

    ```{r}
    dredged <- dredge(model_fitRan, rank = 'BIC')
    head(dredged)
    ```

    **Explanation:**

-   This indicates to us the inclusion of `(1|Partner)` was a smart decision and we should proceed with our assessment!

    **Normality of Residuals**

    ```{r}
    gf_histogram(~resid(model_fitRan), data = data1, bins = 6)|>
      gf_labs(title = "Histogram: Normality of residuals", x = "residuals", y = "count")
    ```

    **Explanation:**

-   Normality of residuals refers to the assumption that the residuals of a statistical model are normally distributed. Residuals are the differences between the observed values and the predicted values from the model. If the residuals are normally distributed, it means that the model is capturing the underlying pattern in the data and that the model is a good fit for the data.

    A good indication of normality of residuals is when the histogram of residuals is approximately bell-shaped and symmetric. **= FAILED**

    **With random effects ACF TEST**

    ```{r}
    s245::gf_acf(~resid(model_fitRan))|>
    gf_lims(y=c(-1,1))|>
       gf_labs(subtitle = "Assement1", title = "Independence Of Residuals", )
    ```

    **Explanation:**

-   The **`gf_acf`** function takes a time series object as input and creates a plot of the autocorrelation function (ACF) for the series. The ACF plot shows the correlation of a time series with its own lagged values. In other words, it shows how correlated a time series is with its own past values at different lags.

    The plot typically has the lag on the x-axis and the correlation coefficient on the y-axis. The plot also includes confidence bands to help identify statistically significant correlations.

-   Besides initial lag exceeding dashed areas, this assessment passes with good conditions

    **Scaled Residuals**

    ```{r}
    food_sim <- simulateResiduals(model_fitRan)
    gf_point(food_sim$scaledResiduals ~ fitted(model_fitRan),
    alpha = 0.2) %>%
    gf_labs(
    y = 'Scaled Residuals')
    ```

    **Results:** Mean variance relationship looks great because we see a vertical relationship with minimal clustering in it.

    **Prediction Plot**

    ```{r}
    ggpredict(model_fitRan, 
              c("Increased_Knowledge", "Increased_Yield"), 
              type = 're') |> 
      plot() 
    ```

    **Results:** I feel that this prediction plot demonstrates a better interaction than what we had in the previous test because it is showing us what we want to see in a interaction predictors. This is also a individual Average rather than population due to the dredge suggesting that a individual average is better.

    ## Conclusion

-   **restated question:** do the programs affect the individual's intellectuality and yield where it makes sure that there is less food deprivation occurring more often. (is there a association between food_ran_out and increased_knowledge/increased_yield?)

-   **Answered based on previous test:** without random effects: can not be concluded from the results of the previous test to say that increased knowledge and increased yield resulted in less food deprivation if anything it could not be trusted for anything.

-   **Answered based on this test:** I can say more confidently that i can see that the interaction is giving us some insight that they might lead us to think that food deprivation occurs less. I would stick to these results because it gives a more affirmative result with appropriate measures followed. The prediction plot is completely different and is showing a interaction.

    ## Limitations of study

-   **Time:** Being a junior with a double major, it was difficult to devote more time into this study because of other commitments.

-   **Knowledge:** I wish i knew more technical knowledge to do more assessments and an evaluations of the model I had!

    ## Implications of the study

-   The extractions that we can make from this study are detrimental because it gives us justification to further expand our programs for food security. It is also necessary to establish initial steps to say that we can further expand or maintain these programs in the country because they are beneficial.

    ## Next step?

-   deduce which variables are causing most influence on my response variable and make a concrete study on that independently.

### potato
